Community-Conditioned Language Models

### 2020-11-17 paper/paper.tex
use cleveref everywhere

We trained the models until the validation loss stopped decreasing 
for two epochs in a row,
and used the weights from the epoch with the smallest validation loss for testing.
In general, transformer models trained for about half as long as the LSTM models (Table~\ref{tab:best-epoch}).

@@

As one would expect, 
the conditioned models have lower perplexity
than their respective unconditioned baseline models,
but this effect is different in magnitude
for models with different community embedding depths .
For the LSTM encoder, 
the best model ($l_c=1$) concatenates the community embedding after the first encoder layer,
with two more LSTM layers after that.
The best transformer model ($l_c=3$), on the other hand,
incorporates the community information last,
after all the sequence encoder layers
(Table \ref{tab:model-ppls}).

@@

We also note that the conditioned models perform differently across different communities---%
for some communities, perplexity of the conditioned models is a lot lower,\bn{is this ok? took ``test'' out but made clear in the caption that it's the test set}%\jp{and training?}
whereas for others it is the same or even slightly higher (Table~\ref{tab:model-ppls}).\bn{talk about this in terms of mutual information instead?}
As an informal observation, we note 
that across all the models we tested, 
communities where conditioning has the least effect
tend to be more organized around more ``general interest'' topics,
such as /r/relationships, where the subject matter is relevant to a broad
range of people. \bn{I really mean this as a quality of the topic, rather than the community itself, although the two are probably correlated}% \jp{Is there a way to quantify general interest topics? Community size in terms of number of users?}
Conditioning the model on community has the most benefit
for narrower special-interest subreddits such as those organized
around a certain videogame, sports team, or subculture.
This makes sense intuitively, since communities with more niche topics
would tend to have more specialized vocabulary and speech patterns.\bn{i think it's ok---i've seen this broader sense used in academic papers}%\jp{can you say speech for written language?}

@@

As a qualitative assesment of the community embeddings, 
we manually identified topic categories into which at least three
of the selected communities fall.
We plotted the two-component PCA projection of the embeddings,
annotated with these categories (Figure~\ref{fig:comm-pca}).
We see that there is some evidence of clustering according to
the topic categories we identified,
suggesting that the embedding space learned by the model is meaningful.

@@

As a quantitative assessment, we compare the CCLM-learned community
embedding with the community embedding created by \citet{Kumar2018},
who generated them from the message-community co-occurence matrix,
using data which spans January 2014 to April 2017. We refer the reader
to \citet{Kumar2018}\jp{I actually cannot find in their paper what
  they do.} for details, but the important point is that no linguistic
information is used to create these embeddings: they only depend on
whether and how much a given user participates in a given
community. In sharp contrasts, CCLM community embeddings depend in no
way on which user is the author of any given message: we only use the
contents of messages, not authorship data. 
%
We correlate the pairwise cosine similarity of the
CCLM embeddings with the same metric on the membership embedding,
restricted to the 46 communities chosen for our study
(Figure~\ref{fig:pairwise-comm-sim}).

\begin{figure}
  \pairwisecommsim{lstm-3-1} 
  \caption{Pairwise cosine similarity between community vectors for 
    the membership-based embedding (x-axis) and
  the LSTM CCLM embedding with $c=1$ (y-axis). 
  Pairs from the same category are colored according to Figure~\ref{fig:comm-pca}.
  We also mark pairs from different (\texttimes) and pairs where one or both
  communities belong to the ``Other'' category (\textbigcircle).
  }
  \label{fig:pairwise-comm-sim}
\end{figure}

@@

However we observe that the CCLM perplexity is only weakly correlated with the
indiscernibility to the corresponding latent LMCC
(\cref{fig:cclm_lmcc_ppl}).
\jp{actually, if we had many more communities, then the lines could be ``more blurred'' between various communities, and we may have a stronger correlation here.}
\begin{figure}
  \pplscatter{lstm-3-1_cclm}{lstm-3-1_lmcc}
  \caption{%
    Mean linguistic perplexity (x-axis) and community indiscernibility
    (y-axis) per community (Pearson's r = \num{0.30}, p =
    \num{0.0444}).  Shown here for the LSTM model with $c = 1$, but
    other models exhibit similar behavior.  }
  \label{fig:cclm_lmcc_ppl}
\end{figure}

>>>>

We trained the models until the validation loss stopped decreasing 
for two epochs in a row,
and used the weights from the epoch with the smallest validation loss for testing.
In general, transformer models trained for about half as long as the LSTM models (\cref{tab:best-epoch}).

@@

As one would expect, 
the conditioned models have lower perplexity
than their respective unconditioned baseline models,
but this effect is different in magnitude
for models with different community embedding depths .
For the LSTM encoder, 
the best model ($l_c=1$) concatenates the community embedding after the first encoder layer,
with two more LSTM layers after that.
The best transformer model ($l_c=3$), on the other hand,
incorporates the community information last,
after all the sequence encoder layers
(\cref{tab:model-ppls}).

@@

We also note that the conditioned models perform differently across different communities---%
for some communities, perplexity of the conditioned models is a lot lower,\bn{is this ok? took ``test'' out but made clear in the caption that it's the test set}%\jp{and training?}
whereas for others it is the same or even slightly higher (\cref{tab:model-ppls}).\bn{talk about this in terms of mutual information instead?}
As an informal observation, we note 
that across all the models we tested, 
communities where conditioning has the least effect
tend to be more organized around more ``general interest'' topics,
such as /r/relationships, where the subject matter is relevant to a broad
range of people. \bn{I really mean this as a quality of the topic, rather than the community itself, although the two are probably correlated}% \jp{Is there a way to quantify general interest topics? Community size in terms of number of users?}
Conditioning the model on community has the most benefit
for narrower special-interest subreddits such as those organized
around a certain videogame, sports team, or subculture.
This makes sense intuitively, since communities with more niche topics
would tend to have more specialized vocabulary and speech patterns.\bn{i think it's ok---i've seen this broader sense used in academic papers}%\jp{can you say speech for written language?}

@@

As a qualitative assesment of the community embeddings, 
we manually identified topic categories into which at least three
of the selected communities fall.
We plotted the two-component PCA projection of the embeddings,
annotated with these categories (\cref{fig:comm-pca}).
We see that there is some evidence of clustering according to
the topic categories we identified,
suggesting that the embedding space learned by the model is meaningful.

@@

As a quantitative assessment, we compare the CCLM-learned community
embedding with the community embedding created by \citet{Kumar2018},
who generated them from the message-community co-occurence matrix,
using data which spans January 2014 to April 2017. We refer the reader
to \citet{Kumar2018}\jp{I actually cannot find in their paper what
  they do.} for details, but the important point is that no linguistic
information is used to create these embeddings: they only depend on
whether and how much a given user participates in a given
community. In sharp contrasts, CCLM community embeddings depend in no
way on which user is the author of any given message: we only use the
contents of messages, not authorship data. 
%
We correlate the pairwise cosine similarity of the
CCLM embeddings with the same metric on the membership embedding,
restricted to the 46 communities chosen for our study
(\cref{fig:pairwise-comm-sim}).

\begin{figure}
  \pairwisecommsim{lstm-3-1} 
  \caption{Pairwise cosine similarity between community vectors for 
    the membership-based embedding (x-axis) and
  the LSTM CCLM embedding with $c=1$ (y-axis). 
  Pairs from the same category are colored according to \cref{fig:comm-pca}.
  We also mark pairs from different (\texttimes) and pairs where one or both
  communities belong to the ``Other'' category (\textbigcircle).
  }
  \label{fig:pairwise-comm-sim}
\end{figure}

@@

However we observe that the CCLM perplexity is only weakly correlated with the
indiscernibility of the corresponding latent LMCC(\cref{fig:cclm_lmcc_ppl}).

\jp{actually, if we had many more communities, then the lines could be ``more blurred'' between various communities, and we may have a stronger correlation here.}
\begin{figure}
  \pplscatter{lstm-3-1_cclm}{lstm-3-1_lmcc}
  \caption{%
    Mean linguistic perplexity (x-axis) and community indiscernibility
    (y-axis) per community (Pearson's r = \num{0.30}, p =
    \num{0.0444}).  Shown here for the LSTM model with $c = 1$, but
    other models exhibit similar behavior.  }
  \label{fig:cclm_lmcc_ppl}
\end{figure}
